# Introduction {.unnumbered}

The focus of this course is int the programming and basic techniques for
inference that are usually applied in data science. We start by reviewing and
enforcing programming skills. Then we will use the database of entomological
data practice and build the required bases for more structured tools like
bootstrap or Jackknife cuts.

@fig-airquality further explores the impact of temperature on ozone level.

```{r}
#| label: fig-airquality
#| fig-cap: Temperature and ozone level.
#| warning: false

library(ggplot2)

ggplot(airquality, aes(Temp, Ozone)) + 
  geom_point() + 
  geom_smooth(method = "loess") + 
  theme_minimal()
  
```

We also want to cover the fundamentals of a common Data Analyst hast to 
know according to the DataCamp's rubrics at the current date (March-01, 2023).

# Rubrics for a Data Analyst (Associate and professional){.unnumbered}

::: {.panel-tabset}
## Associate
```{r}
header <- c("Topic", "Competency","Sufficient","Insufficient")
topics <- c(
  "Data Validation",
  "Data Visualization",
  "Communication"
  )
competnecy <- c(
  "Assess data quality and perform validation tasks",
  "Create data visualizations to demonstrate the characteristics of data and 
  represent relationships between features",
  "Presents data concepts to small, diverse audiences"
) 
sufficeint <- c(
  "Has validated all variables against provided criteria and where necessary has
  performed cleaning tasks to result in analysis-ready data",
  "Has created at least two different visualizations of single variables 
  (e.g. histogram, bar  chart, single boxplot) 
  \n
  Has created at least one visualization including two or more variables 
  (e.g. scatterplot, filled bar chart, multiple boxplots)
  \n
  Has used visualizations that support the findings being presented",
  "For each analysis step, 
  has explained their findings and/or the reasoning for selecting approaches"
)
insufficent <- c(
  "Has not conducted all the required checks and/or has not cleaned the data. 
  May have removed data rather than performed cleaning tasks
  ",
  "
    Has used the same visualization throughout
    \n
    Has not included graphics to represent single variables and relationships
    \n
    Has not used visualizations that support the findings being presented
  ",
  "
    Has not provided a summary for each step 
    (data validation, exploratory analysis)
  "
)

df_rubric_da_associate <- 
  data.frame(
    Topics = topics,
    Competnecy = competnecy,
    Sufficeint = sufficeint,
    Insufficent = insufficent
  )
knitr::kable(df_rubric_da_associate, format = "html")

```
## Professional

```{r}
header <- c("Topic", "Competency","Sufficient","Insufficient")
topics <- c(
  "Data Validation",
  "Data Visualization",
  "Business Focus",
  "Business Metrics",
  "Communication"
  )
competency <- c(
  "
    Assess data quality and perform validation tasks
  ",
  "
    Create data visualizations to demonstrate the characteristics of data and 
    represent relationships between features  ",
  "
    Collects relevant information, detects patterns, observes and interprets 
    data
  ",
  "
    Benchmarks, monitors, and evaluates business processes
  ",
  "
    Employs multiple tactics (written and verbal) to communicate to business 
    leaders
  "
)
sufficient <- c(
  "
    Has validated all variables and where necessary has performed cleaning 
    tasks to result in analysis-ready data
  ",
  "
    Has created at least two different visualizations of single variables 
    (e.g. histogram, bar chart, single boxplot)
    \n
    Has created at least one visualization including two or more variables 
    (e.g. scatterplot, filled barchart, multiple boxplots)
    \n
    Has used visualizations that support the findings being presented
  ",
  "
    Has described at least one of the business goals of the project
    \n
    Has explained how their work has addressed the business problem
    \n
    Has provided at least one recommendation for future action to be taken 
    based on the outcome of the work done
  ",
  "
    Has defined a metric that can be used by the business in the future to 
    measure success in solving the problem
    
    Has evaluated the metric using the existing data to provide a baseline 
    measure for the problem
  ",
  "
    For each analysis step, has provided a written explanation of their findings
    and/or reasoning for selecting approaches
    \n
    Has delivered a verbal presentation addressing the business goals, outcomes
    and recommendations
  "
)
insufficient <- c(
  "
    Has not conducted all the required checks and/or has not cleaned the data. 
    May have removed data rather than performed cleaning tasks
  ",
  "
    Has used the same visualization throughout
    \n
    Has not included graphics to represent single variables and relationships
    \n
    Has not used visualizations that support the findings being presented
  ",
  "
    Has not identified any business goals
    \n
    Has not explained how their work has addressed the business problem
    \n
    Has not provided any recommendations for future actions
  ",
  "
    Has not identified a metric to compare the model performance to the business
    problem or has not shown the metric with the current data
  ",
  "
    Has not provided a written summary for each step
    \n
    Has not delivered a verbal presentation
  "
)

df_rubric_da_professional <- 
  data.frame(
    Topics = topics,
    Competency = competency,
    Sufficient = sufficient,
    Insufficient = insufficient
  )
knitr::kable(df_rubric_da_professional, format = "html")

```
:::



## The tidyverse

We need to install a R package. The majority of the packages that we will use
are part of the so-called tidyverse package. The packages in the tidyverse share
a common philosophy of data and R programming, and are designed to work together
naturally.

You can install the complete tidyverse with the line of code:

```{r}
  # install.packages("tidyverse")
```

then we can use it by loading in the preamble section with

```{r}
  library(tidyverse)
```

see <https://www.tidyverse.org/> documentation.
