# Introduction {.unnumbered}

The focus of this course is int the programming and basic techniques for
inference that are usually applied in data science. We start by reviewing and
enforcing programming skills. Then we will use the database of entomological
data practice and build the required bases for more structured tools like
bootstrap or Jackknife cuts.

@fig-airquality further explores the impact of temperature on ozone level.

```{r}
#| label: fig-airquality
#| fig-cap: Temperature and ozone level.
#| warning: false

library(ggplot2)

ggplot(airquality, aes(Temp, Ozone)) + 
  geom_point() + 
  geom_smooth(method = "loess") + 
  theme_minimal()
  
```

## The tidyverse

We need to install a R package. The majority of the packages that we will use
are part of the so-called tidyverse package. The packages in the tidyverse share
a common philosophy of data and R programming, and are designed to work together
naturally.

You can install the complete tidyverse with the line of code:

```{r}
  install.packages("tidyverse")
```

then we can use it by loading in the preamble section with

```{r}
  library(tidyverse)
```

see <https://www.tidyverse.org/> documentation.
