---
title: "Visualizing Covid-19"
output: html_notebook
---
#  From epidemic to pandemic¶
In December 2019, COVID-19 coronavirus was first identified in the Wuhan region
of China. By March 11, 2020, the World Health Organization (WHO) categorized the
COVID-19 outbreak as a pandemic. A lot has happened in the months in between
with major outbreaks in Iran, South Korea, and Italy.

We know that COVID-19 spreads through respiratory droplets, such as through
coughing, sneezing, or speaking. But, how quickly did the virus spread
across the globe? And, can we see any effect from country-wide policies, like
shutdowns and quarantines?

Fortunately, organizations around the world have been collecting data so that
governments can monitor and learn from this pandemic. Notably, the Johns
Hopkins University Center for Systems Science and Engineering created a publicly
available data repository to consolidate this data from sources like the WHO,
the Centers for Disease Control and Prevention (CDC), and the Ministry of
Health from multiple countries.

In this notebook, you will visualize COVID-19 data from the first several weeks
of the outbreak to see at what point this virus became a global pandemic.

```{r}
library(readr)
library(ggplot2)
library("dplyr")
library(coronavirus)
update_dataset()
# Read datasets/confirmed_cases_worldwide.csv into confirmed_cases_worldwide
path <- getwd()
path <- file.path(
  path,
  "/covid19-visualization-Project",
  "datasets",
  "confirmed_cases_worldwide.csv"
)
confirmed_cases_worldwide <- read_csv(path)
# See the result
confirmed_cases_worldwide



```
# Confirmed cases throughout the world

The table above shows the cumulative confirmed cases of COVID-19 worldwide by
date. Just reading numbers in a table makes it hard to get a sense of the
scale and growth of the outbreak. Let's draw a line plot to visualize the
confirmed cases worldwide.

```{r }
# Draw a line plot of cumulative cases vs. date
# Label the y-axis
confirmed_cases_worldwide %>%
    ggplot(
        aes(
            x=date,
            y=cum_cases
        )
    ) +  geom_line() + ylab("Cumulative confirmed cases")


```
# China compared to the rest of the world

The y-axis in that plot is pretty scary, with the total number of confirmed
cases around the world approaching 200,000. Beyond that, some weird things
are happening: there is an odd jump in mid February, then the rate of new
cases slows down for a while, then speeds up again in March. We need to dig
deeper to see what is happening.

Early on in the outbreak, the COVID-19 cases were primarily centered in China.
Let's plot confirmed COVID-19 cases in China and the rest of the world
separately to see if it gives us any insight.

We'll build on this plot in future tasks. One thing that will be important for
the following tasks is that you add aesthetics within the line geometry of your
ggplot, rather than making them global aesthetics.

```{r}
# Read in datasets/confirmed_cases_china_vs_world.csv
path <- file.path(
  wd_path,
  "/covid19-visualization-Project",
  "datasets",
  "confirmed_cases_china_vs_world.csv"
)
confirmed_cases_china_vs_world <- read_csv(path)

# See the result
glimpse(confirmed_cases_china_vs_world)
# Draw a line plot of cumulative cases vs. date, colored by is_china
# Define aesthetics within the line geom
plt_cum_confirmed_cases_china_vs_world <-
  ggplot(
    data = confirmed_cases_china_vs_world) +
  geom_line(
    aes(
      x = date,
      y = cum_cases,
      color = is_china
    )
  ) +
  ylab("Cumulative confirmed cases")

# See the plot
plt_cum_confirmed_cases_china_vs_world
```
# Let's annotate!

Wow! The two lines have very different shapes. In February, the majority of
cases were in China. That changed in March when it really became a global
outbreak: around March 14, the total number of cases outside China overtook the
cases inside China. This was days after the WHO declared a pandemic.

There were a couple of other landmark events that happened during the outbreak.
For example, the huge jump in the China line on February 13, 2020 wasn't just a
bad day regarding the outbreak; China changed the way it reported figures on
that day (CT scans were accepted as evidence for COVID-19, rather than only
lab tests).

By annotating events like this, we can better interpret changes in the plot.
```{r}
who_events <- tribble(
    ~ date, ~ event,
    "2020-01-30", "Global health\nemergency declared",
    "2020-03-11", "Pandemic\ndeclared",
    "2020-02-13", "China reporting\nchange"
) %>%
    mutate(date = as.Date(date))
# Using who_events, add vertical dashed lines with an xintercept at date
# and text at date, labeled by event, and at 100000 on the y-axis

plt_cum_confirmed_cases_china_vs_world +
  geom_vline(data= who_events,
    aes(xintercept=date)
  ) +
  geom_text(data=who_events,
    aes(x=date, label=event, y=100000)
  )
```
#  Adding a trend line to China
When trying to assess how big future problems are going to be, we need a measure
of how fast the number of cases is growing. A good starting point is to see
if the cases are growing faster or slower than linearly.

There is a clear surge of cases around February 13, 2020, with the reporting
change in China. However, a couple of days after, the growth of cases in
China slows down. How can we describe COVID-19's growth in China after
February 15, 2020?

```{r}
# Filter for China, from Feb 15
china_after_feb15 <- confirmed_cases_china_vs_world %>%
  filter(date > "2020-02-15")

# Using china_after_feb15, draw a line plot cum_cases vs. date
# Add a smooth trend line using linear regression, no error bars
ggplot(china_after_feb15,
       aes(
           x = date,
           y = cum_cases
       )
) +
    geom_line() +
    geom_smooth(method = lm, se = FALSE) +
    ylab("Cumulative confirmed cases")
```
# And the rest of the world?

From the plot above, the growth rate in China is slower than linear. That's
great news because it indicates China has at least somewhat contained the
virus in late February and early March.

How does the rest of the world compare to linear growth?

```{r}
not_china <- confirmed_cases_china_vs_world %>%
  filter( > "2020-02-15")

# Using china_after_feb15, draw a line plot cum_cases vs. date
# Add a smooth trend line using linear regression, no error bars
ggplot(china_after_feb15,
       aes(
           x = date,
           y = cum_cases
       )
) +
    geom_line() +
    geom_smooth(method = lm, se = FALSE) +
    ylab("Cumulative confirmed cases")
```

From the plot above, we can see a straight line does not fit well at all, and
 the rest of the world is growing much faster than linearly. What if we
added a logarithmic scale to the y-axis?
```{r}
# Modify the plot to use a logarithmic scale on the y-axis
plt_not_china_trend_lin +
    scale_y_log10()
```
# Plotting hardest hit countries as of Mid-March 2020¶
Even though the outbreak was first identified in China, there is only one country from East Asia (South Korea) in the above table. Four of the listed countries (France, Germany, Italy, and Spain) are in Europe and share borders. To get more context, we can plot these countries' confirmed cases over time.

Finally, congratulations on getting to the last step! If you would like to
continue making visualizations or find the hardest hit countries as of today,
you can do your own analyses with the latest data available
<https://github.com/RamiKrispin/coronavirus>.
